---
title: "data manipulation"
output: github_document
---

```{r}
library(tidyverse)
```
 ## load in the FAS litter data
 
```{r}
litters_df = read_csv("./data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)
```
 

## `select`

choose some columns but not others.

```{r}
select(litters_df, group, litter_number)
```

```{r}
select(litters_df, group, gd0_weight:gd_of_birth)
```

```{r}
select(litters_df, -litter_number)
```

renaming columns 

```{r}
select(litters_df, GROUP = group, LITTer_NuMBEr = litter_number)
```

also you could use `renmae` to rename without selecting columns!

```{r}
rename(litters_df, GROUP = group, LITTer_NuMBEr = litter_number)
```

Select helpers

choose variables with same begining
```{r}
select(litters_df, starts_with("gd"))
```

!!! if you want to keep litter_number at first column, and keep everything else the same, you can use `everything()` like this!
```{r}
select(litters_df, litter_number, everything())
```

also, the `relocate()` function could also relocate a column to the beginning!

```{r}
relocate(litters_df, litter_number)
```



