---
title: "data manipulation"
output: github_document
---

```{r}
library(tidyverse)
```
 ## load in the FAS litter data
 
```{r}
litters_df = read_csv("./data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)
```
 

## `select`

choose some columns but not others.

```{r}
select(litters_df, group, litter_number)
```

```{r}
select(litters_df, group, gd0_weight:gd_of_birth)
```

```{r}
select(litters_df, -litter_number)
```

renaming columns 

```{r}
select(litters_df, GROUP = group, LITTer_NuMBEr = litter_number)
```

also you could use `renmae` to rename without selecting columns!

```{r}
rename(litters_df, GROUP = group, LITTer_NuMBEr = litter_number)
```

Select helpers

choose variables with same begining
```{r}
select(litters_df, starts_with("gd"))
```

!!! if you want to keep litter_number at first column, and keep everything else the same, you can use `everything()` like this!
```{r}
select(litters_df, litter_number, everything())
```

also, the `relocate()` function could also relocate a column to the beginning!

```{r}
relocate(litters_df, litter_number)
```

## `filter`

```{r}
filter(litters_df, gd0_weight < 22)

filter(litters_df, gd0_weight >= 22)
```

note: use **==**
note: the `!()` means **not equal to**
```{r}
filter(litters_df, gd_of_birth == 20)

filter(litters_df, !(gd_of_birth == 20))
```

```{r}
filter(litters_df, gd_of_birth == 20, gd0_weight >= 22)
```

note: use **==**
```{r}
filter(litters_df, group == "Mod8")
```

note: `%in%`
```{r}
filter(litters_df, group %in% c("Con7", "Mod8"))
```

## `mutate`

note: `str_to_lower()` can change the character into lower cases.
```{r}
mutate(
  litters_df, 
  wt_gain = gd18_weight - gd0_weight,
  group = str_to_lower(group)
  )
```

## `arrange`

```{r}
arrange(litters_df, pups_born_alive)
```

