---
title: "data import"
output: github_document
---

```{r}
library(tidyverse)
```

## read in some data: using relative path (will still work on other people's computer)

read in the litters dataset.

`read_csv` is a function in readr

`clean_names` function did not contains in R or tidyverse, it's in `janitor` package, and we use `janitor::` to realize this without loading all the janitor package.

```{r}
litters_df = read_csv("./data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)
```

## take a look at the data

printing in the console
```{r}
litters_df
```

```{r}
head(litters_df)
tail(litters_df)
```
# we only use one function `skim` in `skimr` package, so we do not load `library(skimr)`. we use `skimr::skim`
```{r}
skimr::skim(litters_df)
```
`view()` function is a useful function, but not usually put it in a code chunk!
```{r}
view(litters_df)
```

## options to read_csv

check out `?read_csv()` for more information

1) skip first 10 rows
2) clear the null data in various formats
```{r}
litters_df = read_csv("./data/FAS_litters.csv", skip = 10, col_names = FALSE)
litters_df = read_csv("./data/FAS_litters.csv", na = c("", "NA", ".", 999))
```

## other file formats

read in excel file.

note: in terms of loading difference packages, you can do it here, or at the begining with `tidyverse`
note: names already clean, so we do not need to clean names!

check `?read_excel`

special options for excel data: choose **sheet** and choose **range** you want to read.

```{r}
library(readxl)
mlb_df = read_excel("./data/mlb11.xlsx")
mlb_df
view(mlb_df)
```

read in SAS file.

need to cleann names.
```{r}
library(haven)
pulse_df = read_sas("./data/public_pulse_data.sas7bdat")
pulse_df
```


## comparision with Base R

what about `read.csv` ...?
`read.csv` is based in the R, and did not to load any package.
but NEVER use this!!!

```{r}
litter_base = read.csv("./data/FAS_litters.csv")
```

the output print is really hard to read!

## exporting data

export the mlb sub_table

```{r}
mlb_df
write_csv(mlb_df, "./data/mlb_subtable.csv")
```

